## Problem 1
### a
```{r}
data = c(1.77, -0.23, 2.76, 3.80, 3.47, 56.75, -1.34, 4.24, 
         -2.44, 3.29, 3.71, -2.40, 4.53, -0.07, -1.05, -13.87, 
         -2.53, -1.75, 0.27, 43.21 )
likelihood = function(x){
  -sum( log(pi*(1+(data-x)^2)) )
}
plot(x = seq(-6,6,0.01), y = sapply(seq(-6,6,0.01), FUN = likelihood), 'l',
     main = 'Cauchy Log-likelihood Function', xlab = 'x', ylab = 'log-likelihood')
dev1likelihood = function(x){
  sum( (2*(data-x)) / (1+(data-x)^2) )
}
dev2likelihood = function(x){
  sum( (2*(data-x)^2-2) / (1+(data-x)^2)^2 )
}
newton_raphson = function(start_point = 2, epsilon = 0.001, max_iter = 200){
  count = 1
  x1 = start_point
  x2 = x1 - dev1likelihood(x1) / dev2likelihood(x1)
  while((abs(x1-x2)>epsilon) & (count < max_iter)){
    x1 = x2
    x2 = x1 - dev1likelihood(x1) / dev2likelihood(x1)
    count = count + 1
  }
  return(c(x2, count))
}
newton_raphson() 
start_points = c(-11, -1, 0, 1.5, 4, 4.7, 7, 8, 38)
sapply(start_points, FUN = newton_raphson)
mean_start_point = mean(start_points)
newton_raphson(mean_start_point)
```
### b
``` {r}
bisection = function(from = -1, to = 1, epsilon = 0.001){
  if(dev1likelihood(from)*dev1likelihood(to)>=0) return('Fail to find solution.')
  x1 = from
  x2 = to
  while(abs(x1-x2)>epsilon){
    y = sapply(c(x1,(x1+x2)/2,x2), FUN = dev1likelihood)
    if(y[1]*y[2] < 0) x2 = (x1+x2)/2
    else if(y[3]*y[2] < 0) x1 = (x1+x2)/2
    else return('Fail to find solution.')
  }
  return((x1+x2)/2)
}
bisection()
bisection(from = 0, to = 2)
```

## Problem 2
### a
```{r}
data = c(3.91, 4.85, 2.28, 4.06, 3.70, 4.04, 5.46, 3.53,
         2.28, 1.96, 2.53, 3.88, 2.22, 3.47, 4.82, 2.46, 
         2.99, 2.54, 0.52, 2.50)
likelihood2 = function(theta){
  sum( log( (1 - cos(data - theta)) / (2*pi) ) )
}
plot(x = seq(-pi,pi,0.01), y = sapply(seq(-pi,pi,0.01), FUN = likelihood2), 'l',
     main = 'Log-likelihood Function', xlab = 'x', ylab = 'log-likelihood')

```

### b
$$
EX = \pi+\sin{\theta} \\
\theta = \arcsin{(EX-\pi)} \\
\hat{\theta} = \arcsin{(\bar{X}-\pi)}
$$

### c
```{r}
moment_est = asin(mean(data)-pi)
dev1likelihood = function(x){
  sum( -sin(data-x) / (1-cos(data-x)) )
}
dev2likelihood = function(x){
  sum( 1 / (cos(data-x)-1) )
}
newton_raphson = function(start_point = 2, epsilon = 0.001, max_iter = 300){
  count = 1
  x1 = start_point
  x2 = x1 - dev1likelihood(x1) / dev2likelihood(x1)
  while((abs(x1-x2) > epsilon) && (count < max_iter)){
    x1 = x2
    x2 = x1 - dev1likelihood(x1) / dev2likelihood(x1)
    count = count + 1
  }
  return(c(x2, count, start_point))
}
newton_raphson(moment_est)
```

### d
```{r}
out = sapply(seq(-pi, pi, length.out = 200), FUN = newton_raphson)
data.frame(out, row.names = c("result", "iter_step", "init_point"))
```


### e 
The two near point is 0.4893938 and 0.5209676, their corresponding solutions are -0.0119720 and 0.5219336.

## Problem 3
### a
For those uncensored, it is trivial that the p.d.f is $f(t_i)$. For those censored, the probability at the censored time is $S(t_i)=P(T>t_i)$.
$$
\log L = \sum_i\left[(1-\omega_i)\log S(t_i) + \omega_i \log f(t_i)\right] \\
= \sum_i\left[(1-\omega_i)(-\Lambda(t_i)\exp(x^T_i\beta)) + \omega_i (\log f(t_i)+x^T_i\beta-\Lambda(t_i)\exp(x^T_i\beta))\right] \\
= \sum_i (\omega_i\log(\mu_i)-\mu_i) + \sum_i\omega_i\log(\frac{\lambda(t_i)}{\Lambda(t_i)})
$$
where $\mu_i = \Lambda(t_i)\exp(x^T_i\beta)$.

### b
```{r}
data = c(6, 6, 6, 6, 7, 9, 10, 10, 11, 13, 
         16, 17, 19, 20, 22, 23, 25, 32, 32,
         34, 35, 1, 1, 2, 2, 3, 4, 4, 5, 5, 
         8, 8, 8, 8, 11, 11, 12, 12, 15, 17, 22, 23)
censor = c(0,1,1,1,1,0,0,
           1,0,1,1,0,0,0,
           1,1,0,0,0,0,0,
           1,1,1,1,1,1,1,
           1,1,1,1,1,1,1,
           1,1,1,1,1,1,1)
treatment = rep(0:1,each=21)
likelihood = function(a, b1, b0){
  lambda = a * data ^ (a-1)
  lambda_cap = data ^ a
  mu = lambda_cap * exp(b0+b1*treatment)
  sum( censor * log(mu * lambda / lambda_cap) - mu )
}
likelihood(1,2,2)
dev1likelihood = function(a, b1, b0){
    c(sum(censor*log(data)-a*data^(a-1)*exp(b0+b1*treatment) + censor/a),
      sum(censor*treatment - data^a*treatment*exp(b0+b1*treatment)),
      sum(censor - data^a*exp(b0+b1*treatment)))
}
dev2likelihood = function(a, b1, b0){
    ao = c(sum(-a*(a-1)*data^(a-2)*exp(b0+b1*treatment) - data^(a-1)*exp(b0+b1*treatment) - censor/a^2),
          sum(-a*data^(a-1)*treatment*exp(b0+b1*treatment)),
          sum(-data^(a-1)*exp(b0+b1*treatment)))
    b1o = c(sum(-a*data^(a-1)*exp(b0+b1*treatment)*treatment),
          sum(-data^a*treatment^2*exp(b0+b1*treatment)),
          sum(-data^a*exp(b0+b1*treatment)*treatment))
    b0o = c(sum(-a*data^(a-1)*exp(b0+b1*treatment)),
          sum(-data^a*treatment*exp(b0+b1*treatment)),
          sum(-data^a*exp(b0+b1*treatment)))
    return(cbind(ao,b1o,b0o))
}
newton_raphson = function(start_point = c(1,1,1), epsilon = 0.01, max_iter = 3000){
  count = 1
  x1 = start_point
  x2 = x1 - solve(dev2likelihood(x1[1],x1[2],x1[3])) %*% dev1likelihood(x1[1],x1[2],x1[3]) 
  while((sum(abs(x1-x2)) > epsilon) && (count < max_iter)){
    x1 = x2
    x2 = x1 - solve(dev2likelihood(x1[1],x1[2],x1[3])) %*% dev1likelihood(x1[1],x1[2],x1[3])
    count = count + 1
  }
  return(c(x2, count))
}
newton_raphson()
```

## Problem 4

The non-linear programming problem is modeled as:
$$
\min_{x_1,x_2} x_2-x_1 \\ 
\text{s.t.}\quad x_1,\ x_2 \ge 0,\\
x_2-x_1 \ge 0,\\
-(x_2+1)e^{-x_2} + (x_1+1)e^{-x_1} = 0.95
$$
Solve by R code:
```{r}
library(Rdonlp2)
obj = function(x){
  x[2]-x[1]
}
par.l = c(0, 0)
par.u = c(Inf, Inf)
constrain1 = function(x){
  - (x[2]+1)*exp(-x[2]) + (x[1]+1)*exp(-x[1])
}
constrain2 = function(x){
  x[2]-x[1]
}
nlin.l = c(0.95, 0)
nlin.u = c(0.95, Inf)
init_point = c(1,4)
ret = donlp2(init_point, obj, par.u=par.u, par.l=par.l, 
             nlin=list(constrain1, constrain2), 
             nlin.u=nlin.u, nlin.l=nlin.l)
ret$par
```
The narrowest possible interval containing 95% of the posterior probability is $[0.04236333,\quad 4.76516825]$.




